Population structure and phenotypic variation of \emph{Sclerotinia sclerotiorum} from dry bean (\emph{Phaseolus vulgaris}) in the United States

### 2017-11-09 doc/manuscript/manuscript.tex
update manuscript to current version

\title{Population structure and phenotypic variation of \emph{Sclerotinia
sclerotiorum} from dry bean in the United States}

@@

\begin{abstract}
The ascomycete pathogen \emph{Sclerotinia sclerotiorum} is a
necrotrophic pathogen on over 400 known host plants, and is the causal
agent of white mold on dry bean. Currently, there are no known cultivars
of dry bean with complete resistance to white mold. For more than 20
years, bean breeders have been using white mold screening nurseries with
natural populations of \emph{S. sclerotiorum} to screen new cultivars
for resistance. It is thus important to know if the genetic diversity in
populations of \emph{S. sclerotiorum} within these nurseries a) reflect
the genetic diversity of the populations in the surrounding region and
b) are stable over time. Furthermore, previous studies have investigated
the correlation between mycelial compatibility groups (MCG) and
multilocus haplotypes (MLH), but none have formally tested these
patterns. We genotyped 366 isolates of \emph{S. sclerotiorum} from
producer fields and white mold screening nurseries surveyed over 10
years in 2003--2012 representing 11 states, Mexico, France, and
Australia at 11 microsatellite loci resulting in 165 MLHs. Populations
were loosely structured over space and time based on analysis of
molecular variance and discriminant analysis of principal components,
but not by cultivar, aggressiveness, or field source. Of all the regions
tested, only Mexico (n=18) shared no MLHs with any other region. Using a
bipartite network-based approach, we found no evidence that the MCGs
accurately represent MLHs. Our study suggests that breeders should
continue to test dry bean lines in several white mold screening
nurseries across the US to account for both the phenotypic and genotypic
variation that exists across regions.
% Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text.
\end{abstract}

@@

\emph{Sclerotinia sclerotiorum} (Lib.) de Bary is an ascomycete plant
pathogen with a worldwide distribution. This is a necrotrophic pathogen
that is primarily homothallic (self-fertilization) and has the ability
to survive for more than five years in soil using melanized survival
structures called sclerotia (Bolton et al., 2006; Sexton et al., 2006).
It causes disease on more than 400 plant species belonging to 75
families (Boland \& Hall, 1994) including crops of major economic
importance such as sunflower (\emph{Helianthus spp.}), soybean
(\emph{Glycine max} L.), canola (\emph{Brassica napa} L., \emph{Brassica
campestris} L.), and dry bean (\emph{Phaseolus vulgaris} L.) (Bolton et
al., 2006). On dry bean, \emph{S. sclerotiorum} is the causal agent of
white mold, a devastating disease that can be yield-limiting in
temperate climates (Steadman, 1983).

Currently, there are no commercially available resistant cultivars of
dry bean (Otto-Hanson et al., 2011). Organized breeding efforts have
used a common-garden approach with white mold screening nurseries in dry
bean production areas across the United States with additional sites in
Mexico, France, and Australia (Steadman et al., 2003, 2004, 2005, 2006;
Otto-Hanson \& Steadman, 2007, 2008; McCoy \& Steadman, 2009). These
white mold screening nurseries use no chemical or cultural treatments
against \emph{S. sclerotiorum} and employ standardized protocols for
screening new cultivars for resistance to white mold. It has previously
been shown that aggressiveness (the severity of disease symptoms on the
host) is significantly different across white mold screening nursery
sites in separate geographic regions (Otto-Hanson et al., 2011). The
genetic structure and mode of reproduction in these populations,
however, is currently unknown.

@@

Nearly all population genetic studies of \emph{S. sclerotiorum} employ a
macroscopic assay to determine mycelial compatibility, the ability for
fungal hyphae from different colonies to appear to grow together without
forming a barrier of dead cells between them (known as a barrage line)
(Leslie, 1993; Sirjusingh \& Kohn, 2001). Mycelial compatibility has
been used as a proxy for vegetative compatibility, a fungal trait
controlled by several independent genes controlling the ability for two
hyphae to fuse and grow as a single unit (Leslie, 1993; Schafer \& Kohn,
2006). Because of the genetic connection to vegetative compatibility,
two isolates that are mycelially compatible were considered clones
(Leslie, 1993); but correlation with genetic markers, such as
microsatellites, have shown mixed results (Ford et al., 1995; Micali \&
Smith, 2003; Jo et al., 2008; Attanayake et al., 2012; Papaioannou \&
Typas, 2014; Lehner et al., 2017).

In this study, we analyze and characterize the genetic and phenotypic
diversity of 366 \emph{S. sclerotiorum} isolates collected between 2003
and 2012 from dry bean cultivars among different geographic locations in
the United States, France, Mexico, and Australia. We wanted to know if
the \emph{S. sclerotiorum} populations from white mold screening
nurseries were representative of the fields within the same region. As
these nurseries were not treated with any chemical or cultural control
of white mold, we hypothesized that these nurseries would represent the
natural population of \emph{S. sclerotiorum}. Furthermore, we wanted to
investigate the potential effect of cultivar on genetic diversity of the
pathogen by assessing three dry bean cultivars with different levels of
resistance, Beryl (great northern bean, susceptible), Bunsi (navy bean,
low susceptibility), and G122 (cranberry bean, partial resistance)
(Otto-Hanson et al., 2011). We additionally wanted to determine
categorical or phenotypic variables that best predicted genetic
structure and if there was correlation between multilocus haplotype and
mycelial compatibility group. Knowing what variables predict genetic
structure can help direct breeding efforts. By investigating these aims,
we hope to effectively describe the population structure of \emph{S.
sclerotiorum} in the USA and make available our database of isolates for
use in future dry bean breeding efforts.

@@

Several of the isolates used for this study were collected as reported
in previous studies using the same methods (Otto-Hanson et al., 2011).
Broadly, isolates were collected from two sources: white mold screening
nurseries (wmn) or producer fields. White mold screening nurseries were
5m x 10m in size and maintained without application of fungicides to
observe natural incidence of white mold. The early nursery plots were
incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of
0:3:2 during the growing season (Steadman et al., 2003).

@@

Sclerotia of \emph{S. sclerotiorum} were collected over several years
from grower fields and/or wmn in 11 states of the United States of
America, Mexico, France, and Australia (Table \ref{tab:isolate-table}).
After collection, sclerotia were stored in Petri plates lined with
filter paper, then stored at \(20~^{\circ}\)F or -\(4~^{\circ}\)C.
Sclerotia were surface-sterilized with 50\% Clorox bleach (at least 6\%
NaOCL, The Clorox Company, Oakland, CA) solution for 3 min, and double
rinsed with ddH\(_2\)O for 3 min. The sterilized sclerotia were then
placed on water agar plates (16g of Bacto agar per liter of ddH\(_2\)O,
BD Diagnostic Systems, Sparks, MD), with four to five sclerotia of each
isolate separated on each plate and stored on the counter top at room
temperature for 5 to 6 days. An 8-mm plug from a 5- or 6-day-old culture
was transferred from the advancing margin of the mycelia onto a plate of
Difco potato dextrose agar (PDA at 39 g/liter of ddH\(_2\)O)
(Otto-Hanson et al., 2011). In combination with the 156 isolates
described previously, we collected 210 isolates for a total of 366
isolates (Otto-Hanson et al., 2011).

@@

MCG was determined as described previously through co-culturing pairs of
2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium (Cubeta et
al., 2001). Incompatibility of different MCGs resulted in formation of a
barrage line accompanied by formation of sclerotia on either side of the
barrage line, indicating the limits of the isolates' growth (Kohn et
al., 1990; Leslie, 1993; Otto-Hanson et al., 2011). Isolates were
compared in a pairwise manner for each site and then representatives
among sites were compared to determine mycelial compatibility groups by
scoring compatible and incompatible interactions (Otto-Hanson et al.,
2011). No MCGs were compatible with any other MCG.

@@

All data processing and analyses were performed in R version 3.4.1 (R
Core Team, 2017) and are openly available and reproducible at
\url{https://github.com/everhartlab/sclerotinia-366/}. Of the 16
microsatellite loci genotyped, five included compound repeats, which
made it challenging to accurately/confidently bin alleles into fragment
sizes expected for each locus based on the described repeat motif. Loci
with compound repeats were removed for the reported statistics. To
ensure the integrity of the results we additionally processed these loci
and included them in concurrent analyses. We assessed the power of our
11 markers by generating a genotype accumulation curve in the R package
\emph{poppr} version 2.5.0, looking for evidence of saturation, which
would indicate that loci were sufficiently sampled to adequately
represent the full set of haplotypes (Arnaud-Hanod et al., 2007; Kamvar
et al., 2015). To avoid including isolates potentially collected from
the same plant (which increases the probability of collecting sclerotia
from the same point of infection more than once), data were
clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that
duplicated genotypes were reduced to a single observation when they were
collected in the same year from the same host cultivar located in the
same source field (wmn or producer)---for subsequent analysis. We
assessed haplotype diversity by calculating Stoddart and Taylor's index
(\emph{G}) (Stoddart \& Taylor, 1988), Shannon's index (\emph{H})
(Shannon, 1948), Simpson's index (\(\lambda\)) (Simpson, 1949), evenness
(\(E_5\)), and the expected number of multilocus haplotypes
(\emph{eMLH}) (Hurlbert, 1971; Heck et al., 1975; Pielou, 1975; Grünwald
et al., 2003). To assess the potential for random mating, we tested for
linkage disequilibrium with the index of association, \(I_A\) and its
standardized version, \(\bar{r}_d\) using 999 permutations (Brown et
al., 1980; Smith et al., 1993; Agapow \& Burt, 2001). Both haplotype
diversity and linkage disequilibrium were calculated in \emph{poppr}
(Kamvar et al., 2014).

@@

We used ANOVA to assess if aggressiveness was significantly different
with respect to Region, MCG, or multilocus haplotype (MLH). To minimize
complications due to small sample sizes, we chose the top 10 MCGs,
representing 56.5\% of the isolates collected, the 10 most abundant MLHs
representing 26.7\% of the isolates, and populations with more than five
isolates. If ANOVA results were significantly different at \(\alpha\) =
0.05, pairwise differences were assessed using Tukey's HSD test
(\(\alpha\) = 0.05) using the \texttt{HSD.test()} function in the
package \emph{agricolae} version 1.2.8 (Mendiburu \& Simon, 2015).

@@

We wanted to evaluate patterns of connectivity between shared multilocus
haplotypes across geographic regions. We first tabulated the multilocus
haplotypes shared between at least two populations (defined as states or
countries) with the \emph{poppr} function \texttt{mlg.crosspop()}
(Kamvar et al., 2014). From these data, we constructed a graph with
populations as nodes and shared haplotypes as edges (connections)
between nodes using the R packages \emph{igraph} (Csardi \& Nepusz,
2006), \emph{dplyr} version 0.7.4 (Wickham et al., 2017), and
\emph{purrr} version 0.2.3 (Henry \& Wickham, 2017). Each node was
weighted by the fraction of shared MLHs. Each edge represented a single
MLH, but because a single MLH could be present in more than one
population, that MLH would have a number of edges equivalent to the
total number of possible connections, calculated as
(\emph{n}*(\emph{n}-1))/2 edges where n represents the number of
populations crossed. Edges were weighted by \(1-P_{sex}\), where
\(P_{sex}\) is the probability of encountering the same haplotype via
two independent meiotic events (Parks \& Werth, 1993; Arnaud-Hanod et
al., 2007). This weighting scheme would thus strengthen the connection
of edges that represented genotypes with a low probability of being
produced via sexual reproduction. We then identified communities (among
the Regions) in the graph using the \texttt{cluster\_optimal()} function
from \emph{igraph} (Csardi \& Nepusz, 2006). The graph was plotted using
the R packages \emph{ggplot2} version 2.2.1 (Wickham, 2009) and
\emph{ggraph} 1.0.0 (Pedersen, 2017). To ensure that we captured the
same community signal, we additionally performed this analysis including
the five polymorphic markers described above.

@@

A total of 366 isolates were collected from 2003 to 2012 (except 2006
and 2011) from diseased dry bean plants in Australia, Mexico, France,
and 11 states in the USA (Table \ref{tab:isolate-table}). We observed a
total of 165 MLHs (215 with 16 loci). Over 50\% of the isolates came
from four states, MI (62), ND (60), WA (59), NE (47). Four regions had
fewer than 10 isolates, Australia (6), WI (2), NY (1), ID (1). We
observed 87 MCGs, the most abundant of which (`MCG 5') was represented
by 73 isolates over 37 MLHs (Fig \ref{biggraph}A,C).

The number of observed alleles per locus ranged from two to 10 with an
average of 6.27 (Table \ref{tab:locus-stats}). Locus 20-3, which
contained only 2 alleles, showed low values of both \emph{h} (0.0533)
and evenness (0.42), indicating that there was one dominant allele
present. Analysis of the haplotype accumulation curve showed no clear
plateau for 11 or 16 loci (Supplemental Information), indicating that we
would likely obtain more multilocus haplotypes if we were to genotype
more loci.

After clone-correction on the hierarchy of Region/Source/Host/Year, we
were left with 318 isolates representing 165 MLHs (Table
\ref{tab:iatab}). The results show that most populations exhibited
relatively high amounts of genotypic diversity, with the exception of MI
and NE, which exhibited lower-than average \(E_5\) values. Mexico had
the lowest average value of gene diversity, \emph{h} and the lowest
\emph{eMLH}, suggesting low overall genetic diversity. Nearly all
populations showed evidence for linkage (Table \ref{tab:iatab}), which
serves as evidence for clonal reproduction or other forms of non-random
mating. The only exceptions are CA (\emph{P} = 0.043) and Australia
(\emph{P} = 0.052). Both of these populations showed only moderate
significance with \(\bar{r}_d\) values of 0.03 and 0.12, respectively.

\begin{longtable}[]{@{}rrllrl@{}}

@@

\end{minipage} & \begin{minipage}[b]{0.15\columnwidth}\raggedright\strut

@@

\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
di-\strut

@@

\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
hexa-\strut

@@

\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
di-\strut

@@

\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
di-\strut

@@

\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
di-\strut

@@

\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
tri-\strut

@@

\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
di-\strut

@@

\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
tetra-\strut

@@

\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
tetra-\strut

@@

\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
tetra-\strut

@@

\begin{figure}
\centering
\includegraphics{../../results/figures/publication/aggressiveness.pdf}
\caption{Strip Plot of aggressiveness by population arranged in
descending order of mean aggressiveness for all populations with N
\textgreater{} 5. White bars represent mean value. Circles represent
individual isolates where filled circles are isolates from white mold
screening nurseries (wmn) and open circles are isolates from producer
fields.}\label{aggressiveness-boxplot}
\end{figure}

>>>>

\title{Population structure and phenotypic variation of \emph{Sclerotinia
sclerotiorum} from dry bean (\emph{Phaseolus vulgaris}) in the United
States}

@@

\begin{abstract}
The ascomycete pathogen \emph{Sclerotinia sclerotiorum} is a
necrotrophic pathogen on over 400 known host plants, and is the causal
agent of white mold on dry bean. Currently, there are no known cultivars
of dry bean with complete resistance to white mold. For more than 20
years, bean breeders have been using white mold screening nurseries with
natural populations of \emph{S. sclerotiorum} to screen new cultivars
for resistance. It is thus important to know if the genetic diversity in
populations of \emph{S. sclerotiorum} within these nurseries a) reflect
the genetic diversity of the populations in the surrounding region and
b) are stable over time. Furthermore, previous studies have investigated
the correlation between mycelial compatibility groups (MCG) and
multilocus haplotypes (MLH), but none have formally tested these
patterns. We genotyped 366 isolates of \emph{S. sclerotiorum} from
producer fields and white mold screening nurseries surveyed over 10
years in 2003--2012 representing 11 states in the United States of
America, Australia, France, and Mexico at 11 microsatellite loci
resulting in 165 MLHs. Populations were loosely structured over space
and time based on analysis of molecular variance and discriminant
analysis of principal components, but not by cultivar, aggressiveness,
or field source. Of all the regions tested, only Mexico (n=18) shared no
MLHs with any other region. Using a bipartite network-based approach, we
found no evidence that the MCGs accurately represent MLHs. Our study
suggests that breeders should continue to test dry bean lines in several
white mold screening nurseries across the US to account for both the
phenotypic and genotypic variation that exists across regions.
% Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text.
\end{abstract}

@@

\emph{Sclerotinia sclerotiorum} (Lib.) de Bary is an ascomycete plant
pathogen with a worldwide distribution (Bolton et al., 2006). This is a
necrotrophic pathogen that is primarily homothallic (self-fertilization)
and has the ability to survive for more than five years in soil using
melanized survival structures called sclerotia (Bolton et al., 2006;
Sexton et al., 2006). It causes disease on more than 400 plant species
belonging to 75 families (Boland \& Hall, 1994) including crops of major
economic importance such as sunflower (\emph{Helianthus spp.}), soybean
(\emph{Glycine max} L.), canola (\emph{Brassica napa} L., \emph{Brassica
campestris} L.), and dry bean (\emph{Phaseolus vulgaris} L.) (Bolton et
al., 2006).

On dry bean, \emph{S. sclerotiorum} is the causal agent of white mold, a
devastating disease that can be yield-limiting in temperate climates
(Steadman, 1983). All above-ground tissues (flowers, stems, leaves,
pods) are susceptible to infection, first appearing as wet lesions with
white mycelial tufts, and then bleaching as the tissue senesces
(Steadman, 1983; Bolton et al., 2006). For many years, white mold has
been the most serious dry bean disease in the Northwestern United States
(Otto-Hanson et al., 2011; Knodel et al., 2012, 2015, 2016). The impact
of white mold on the dry bean industry in the Northwestern United States
alone has been estimated at a loss of 140 kg/ha with just 10\% disease
incidence (Ramasubramaniam et al., 2008).

Currently, there are no commercially available resistant cultivars of
dry bean (Otto-Hanson et al., 2011). Organized breeding efforts have
used a common-garden approach with white mold screening nurseries in dry
bean production areas across the United States with additional sites in
Australia, France, and Mexico (Steadman et al., 2003, 2004, 2005, 2006;
Otto-Hanson \& Steadman, 2007, 2008; McCoy \& Steadman, 2009). These
white mold screening nurseries use no chemical or cultural treatments
against \emph{S. sclerotiorum} and employ standardized protocols for
screening new cultivars for resistance to white mold (Steadman et al.,
2003; Otto-Hanson et al., 2011). These protocols included three
established cultivars used for comparison in the trials: Beryl (great
northern bean, susceptible), Bunsi (aka Ex Rico, navy bean, low
susceptibility), and G122 (cranberry bean, partial resistance) (Tu \&
Beversdorf, 1982; Steadman et al., 2005; Otto-Hanson et al., 2011). It
has previously been shown that aggressiveness (the severity of disease
symptoms on the host) is significantly different across white mold
screening nursery sites in separate geographic regions (Otto-Hanson et
al., 2011). The genetic structure and mode of reproduction in these
populations, however, is currently unknown.

@@

Nearly all population genetic studies of \emph{S. sclerotiorum} employ a
macroscopic assay to determine mycelial compatibility, the ability for
fungal hyphae from different colonies to appear to grow together without
forming a barrier of dead cells between them (known as a barrage line,
Fig. \ref{mcg-fig}B) (Leslie, 1993; Sirjusingh \& Kohn, 2001). Mycelial
compatibility has been used as a proxy for vegetative compatibility, a
fungal trait controlled by several independent genes controlling the
ability for two hyphae to fuse and grow as a single unit (Fig.
\ref{mcg-fig}A) (Leslie, 1993; Schafer \& Kohn, 2006). Because of the
genetic connection to vegetative compatibility, two isolates that are
mycelially compatible were considered clones (Leslie, 1993); but
correlation with genetic markers, such as microsatellites, have shown
mixed results (Ford et al., 1995; Micali \& Smith, 2003; Jo et al.,
2008; Attanayake et al., 2012; Papaioannou \& Typas, 2014; Lehner et
al., 2017). However, a question remains about the relationship between
mycelial compatibility groups and clonal genotypes.

In this study, we analyze and characterize the genetic and phenotypic
diversity of 366 \emph{S. sclerotiorum} isolates collected between 2003
and 2012 from dry bean cultivars among different geographic locations in
the Australia, France, Mexico, and the United States. We wanted to know
if the \emph{S. sclerotiorum} populations from white mold screening
nurseries were representative of the fields within the same region. As
these nurseries were not treated with any chemical or cultural control
of white mold, we hypothesized that these nurseries would represent the
natural population of \emph{S. sclerotiorum}. Furthermore, we wanted to
investigate the potential effect of cultivar on genetic diversity of the
pathogen by assessing three dry bean cultivars with different levels of
resistance, Beryl (great northern bean, susceptible), Bunsi (navy bean,
low susceptibility), and G122 (cranberry bean, partial resistance)
(Otto-Hanson et al., 2011). We additionally wanted to determine
categorical or phenotypic variables that best predicted genetic
structure and if there was correlation between multilocus haplotype and
mycelial compatibility group. Knowing what variables predict genetic
structure can help direct breeding efforts. By investigating these aims,
we will effectively describe the population structure of \emph{S.
sclerotiorum} in the USA and make available our database of isolates for
use in future dry bean breeding efforts.

@@

Several (156) of the isolates used for this study were collected as
reported in previous studies using the same methods (Otto-Hanson et al.,
2011). Broadly, isolates were collected from two sources: white mold
screening nurseries (wmn) or producer fields. White mold screening
nurseries were 5m x 10m in size and maintained without application of
fungicides to observe natural incidence of white mold. The early nursery
plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side
dressing of 0:3:2 during the growing season (Steadman et al., 2003).

@@

Sclerotia of \emph{S. sclerotiorum} were collected over several years
from grower fields and/or wmn in 11 states of the Australia, France,
Mexico, and the United States (Table \ref{tab:isolate-table}). After
collection, sclerotia were stored in Petri plates lined with filter
paper, then stored at \(20~^{\circ}\)F or -\(4~^{\circ}\)C. Sclerotia
were surface-sterilized with 50\% Clorox bleach (at least 6\% NaOCl, The
Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with
ddH\(_2\)O for 3 min. The sterilized sclerotia were then placed on water
agar plates (16g of Bacto agar per liter of ddH\(_2\)O, BD Diagnostic
Systems, Sparks, MD), with four to five sclerotia of each isolate
separated on each plate and stored on the counter top at room
temperature for 5 to 6 days. An 8-mm plug from a 5- or 6-day-old culture
was transferred from the advancing margin of the mycelia onto a plate of
Difco potato dextrose agar (PDA at 39 g/liter of ddH\(_2\)O)
(Otto-Hanson et al., 2011). In combination with the 156 isolates
described previously, we collected 210 isolates for a total of 366
isolates (Otto-Hanson et al., 2011).

@@

MCG was determined as described previously through co-culturing pairs of
2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium (Fig.
\ref{mcg-fig}) (Cubeta et al., 2001). Incompatibility of different MCGs
resulted in formation of a barrage line accompanied by formation of
sclerotia on either side of the barrage line, indicating the limits of
the isolates' growth (Kohn et al., 1990; Leslie, 1993; Otto-Hanson et
al., 2011). Isolates were compared in a pairwise manner for each site
and then representatives among sites were compared to determine mycelial
compatibility groups by scoring compatible and incompatible interactions
(Otto-Hanson et al., 2011). No MCGs were compatible with any other MCG.

@@

All data processing and analyses were performed in a Rocker ``verse''
project container running R version 3.4.2 (Boettiger \& Eddelbuettel,
2017; R Core Team, 2017) and are openly available and reproducible at
\url{https://github.com/everhartlab/sclerotinia-366/}. Of the 16
microsatellite loci genotyped, five included compound repeats, which
made it challenging to accurately/confidently bin alleles into fragment
sizes expected for each locus based on the described repeat motif. Loci
with compound repeats were removed for the reported statistics. To
ensure the integrity of the results we additionally processed these loci
and included them in concurrent analyses. We assessed the power of our
11 markers by generating a genotype accumulation curve in the R package
\emph{poppr} version 2.5.0, looking for evidence of saturation, which
would indicate that loci were sufficiently sampled to adequately
represent the full set of haplotypes (Arnaud-Hanod et al., 2007; Kamvar
et al., 2015). To avoid including isolates potentially collected from
the same plant (which increases the probability of collecting sclerotia
from the same point of infection more than once), data were
clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that
duplicated genotypes were reduced to a single observation when they were
collected in the same year from the same host cultivar located in the
same source field (wmn or producer)---for subsequent analysis. We
assessed haplotype diversity by calculating Stoddart and Taylor's index
(\emph{G}) (Stoddart \& Taylor, 1988), Shannon's index (\emph{H})
(Shannon, 1948), Simpson's index (\(\lambda\)) (Simpson, 1949), evenness
(\(E_5\)), and the expected number of multilocus haplotypes
(\emph{eMLH}) (Hurlbert, 1971; Heck et al., 1975; Pielou, 1975; Grünwald
et al., 2003). To assess the potential for random mating, we tested for
linkage disequilibrium with the index of association, \(I_A\) and its
standardized version, \(\bar{r}_d\) using 999 permutations (Brown et
al., 1980; Smith et al., 1993; Agapow \& Burt, 2001). Both haplotype
diversity and linkage disequilibrium were calculated in \emph{poppr}
(Kamvar et al., 2014).

@@

We used ANOVA to assess if aggressiveness (determined via straw test on
a scale of 1--9) was significantly different with respect to Region,
MCG, or multilocus haplotype (MLH). To minimize complications due to
small sample sizes, we chose the top 10 MCGs, representing 56.5\% of the
isolates collected, the 10 most abundant MLHs representing 26.7\% of the
isolates, and populations with more than five isolates. If ANOVA results
were significantly different at \(\alpha\) = 0.05, pairwise differences
were assessed using Tukey's HSD test (\(\alpha\) = 0.05) using the
\texttt{HSD.test()} function in the package \emph{agricolae} version
1.2.8 (Mendiburu \& Simon, 2015).

@@

We wanted to evaluate patterns of connectivity between shared multilocus
haplotypes across geographic regions. We first tabulated the multilocus
haplotypes shared between at least two populations (defined as states or
countries) with the \emph{poppr} function \texttt{mlg.crosspop()}
(Kamvar et al., 2014). From these data, we constructed a graph with
populations as nodes and shared haplotypes as edges (connections)
between nodes using the R packages \emph{igraph} (Csardi \& Nepusz,
2006), \emph{dplyr} version 0.7.4 (Wickham et al., 2017), and
\emph{purrr} version 0.2.4 (Henry \& Wickham, 2017). Each node was
weighted by the fraction of shared MLHs. Each edge represented a single
MLH, but because a single MLH could be present in more than one
population, that MLH would have a number of edges equivalent to the
total number of possible connections, calculated as
(\emph{n}*(\emph{n}-1))/2 edges where n represents the number of
populations crossed. Edges were weighted by \(1-P_{sex}\), where
\(P_{sex}\) is the probability of encountering the same haplotype via
two independent meiotic events (Parks \& Werth, 1993; Arnaud-Hanod et
al., 2007). This weighting scheme would thus strengthen the connection
of edges that represented genotypes with a low probability of being
produced via sexual reproduction. We then identified communities (among
the Regions) in the graph using the \texttt{cluster\_optimal()} function
from \emph{igraph} (Csardi \& Nepusz, 2006). The graph was plotted using
the R packages \emph{ggplot2} version 2.2.1 (Wickham, 2009) and
\emph{ggraph} 1.0.0 (Pedersen, 2017). To ensure that we captured the
same community signal, we additionally performed this analysis including
the five polymorphic markers described above.

@@

A total of 366 isolates were collected from 2003 to 2012 (except 2006
and 2011) from diseased dry bean plants in 11 states in the United
States as well as Australia, France, and Mexico (Table
\ref{tab:isolate-table}). With the 11 loci used in the analyses (Table
\ref{tab:locus-stats}), we observed a total of 165 MLHs (215 with 16
loci). These 11 loci represented 7 chromosomes in the \emph{S.
sclerotiorum} genome with a minimum distance of 55Kbp between two loci
on the same chromosome. Over 50\% of the isolates came from four states,
MI (62), ND (60), WA (59), NE (47). Four regions had fewer than 10
isolates, Australia (6), WI (2), NY (1), ID (1). We observed 87 MCGs,
the most abundant of which (`MCG 5') was represented by 73 isolates over
37 MLHs (Fig \ref{biggraph}A,C).

The number of observed alleles per locus ranged from two to 10 with an
average of 6.27 (Table \ref{tab:locus-stats}). Locus 20-3, which
contained only 2 alleles, showed low values of both \emph{h} (0.0533)
and evenness (0.42), indicating that there was one dominant allele
present. Analysis of the haplotype accumulation curve showed no clear
plateau for 11 or 16 loci (See section on `Loading Data and Setting
Strata' in the MLG-distribution.md\footnote{Direct link:
  \url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/MLG-distribution.md\#loading-data-and-setting-strata}}
file in the supplemental files (Kamvar et al., 2017)), indicating that
we would likely obtain more multilocus haplotypes if we were to genotype
more loci.

After clone-correction on the hierarchy of Region/Source/Host/Year, a
total of 48 isolates were removed from the data set, resulting in 318
isolates representing 165 MLHs that were used in subsequent analyses
(Table \ref{tab:iatab}). The results showed that, in terms of genotypic
diversity (\emph{H}, \emph{G}, and \(\lambda\)), WA was the most diverse
population with both \emph{G} (54.3) and \(e^H\) (55.3) being close to
the observed number of MLGs (56). This indicated that there are few
duplicated genotypes in WA (Table \ref{tab:iatab}). A more useful metric
to compare populations, however, is \(E_5\), which scales from 0 to 1
where 1 indicates all unique genotypes (Grünwald et al., 2003).
Evaluating by \(E_5\) shows that both MI and NE exhibit lower than
average values, indicating that there are over-represented genotypes in
the popualtions (table \ref{tab:iatab}). When we look at Mexico, we can
observed that it had relatively high values of \(E_5\) and genotypic
diversity, but low richness, as measured by \emph{eMLG}. Moreover,
Mexico had the lowest value for \emph{h}, which is a measure of allelic
diversity. Nearly all populations showed evidence for linkage (Table
\ref{tab:iatab}), which serves as evidence for clonal reproduction or
other forms of non-random mating. The only exceptions are CA (\emph{P} =
0.043) and Australia (\emph{P} = 0.052). Both of these populations
showed only moderate significance with \(\bar{r}_d\) values of 0.03 and
0.12, respectively.

\begin{longtable}[]{@{}rrrlrl@{}}

@@

\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedleft\strut

@@

\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(GT)\strut

@@

\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(TTTTTC)(TTTTTG)(TTTTTC)\strut

@@

\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(GA)\strut

@@

\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(CA)\strut

@@

\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(CA)(CT)\strut

@@

\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(TTA)\strut

@@

\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(GT)GG(GT)\strut

@@

\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(TACA)\strut

@@

\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(TATG)\strut

@@

\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
(TAGA)\strut

@@

\begin{figure}
\centering
\includegraphics{../../results/figures/publication/aggressiveness.pdf}
\caption{Strip plot of aggressiveness by population arranged in
descending order of mean aggressiveness for all populations with N
\textgreater{} 5. White bars represent mean value. Circles represent
individual isolates where filled circles are isolates from white mold
screening nurseries (wmn) and open circles are isolates from producer
fields.}\label{aggressiveness-boxplot}
\end{figure}

The forward-backward selection process of the dbRDA models on
clone-corrected data revealed Year, Region, Host, and MCG to be the
optimal variables for the reduced model, accounting for 45\% of the
total variation. ANOVA showed that the reduced model was significant
with an adjusted R\textsuperscript{2} of 0.0675 (\emph{P} = 0.001).
Assessment of the marginal effects showed that all varibles
significantly explained genetic variation (\(P \leq 0.007\)). We found
that there was multicollinearity when MCG was combined with any other
variable, so repeated the analysis, dropping MCG from the list of
potential predictors. From these results, Year, Region, Host, and
Aggressiveness were found to be optimal, accounting for 17.6\% of the
total variation. ANOVA revealed significant effects with an adjusted
R\textsuperscript{2} of 0.0325 (\emph{P} = 0.001). While the marginal
effect assessment revealed that Year, Region, and Host significantly
explained variation at \emph{P} = 0.001, and Aggressiveness
significantly explained variation at \emph{P} = 0.039. Much of the
variation appeared to be driven by isolates from Mexico and 2005 (Fig.
\ref{rda-biplot}). Variance partitioning of the independent variables
without MCG indicated aggressiveness to be the least influential factor
with 0.1\% contributing to explaining the variation of molecular data,
whereas the combination of variables accounted for 3.3\%.
